---
title: "Ch. 2 Equilibrium Thermodynamics and Kinetics"
output:
  html_document:
    df_print: paged
  word_document: default
date: '2022-06-14'
---
# Equilibrium Thermodynamics
In the real world, systems can exist in several states, *unstable, metastable, and stable*, where an unstable system reacts **spontaneously**, a metastable system requires a minimum **activation energy** to react, and a stable system has reached **equilibrium**. 

### Gibbs Free Energy For a Reaction
To find the total ΔG for a reaction at a given T and P, the total free energies of the reactants are subtracted from that of the reactants.
```{r, message=FALSE}
gibbsReaction <- function(solidDescription)
{
      E.units("J")
      
      ## extract reactant G
      reactant <- data.frame(subcrt(reactant, reactantState, T = Temp, P = Pres))
      nameReactant <- paste("reactant$out.", reactant$species.name, ".G",sep ="")
      ΔGr <- round(eval(parse(text = nameReactant[1]))/1000,3)
      
      ## extract product G
      a <- data.frame(subcrt(productA, T = Temp, P = Pres))
      nameA <- paste("a$out.", str_replace(productA, "[\\+\\-\\*]", "."), ".G",sep ="")
      ΔGa <- round(eval(parse(text = nameA[1]))/1000,3)

      b <- data.frame(subcrt(productB, T = Temp, P = Pres))
      nameB <- paste("b$out.", str_replace(productB, "[\\+\\-\\*]", "."), ".G",sep ="")
      ΔGb <- round(eval(parse(text = nameB[1]))/1000,3)
      
      c <- data.frame(subcrt(productC, T = Temp, P = Pres))
      nameC <- paste("c$out.", str_replace(productC, "[\\+\\-\\*]", "."), ".G",sep ="")
      ΔGc <- round(eval(parse(text = nameC[1]))/1000,3)
      
      ## print chemical formula
      cat("REACTION DESCRIPTION")
      cat("  Reactants ⇌ Products\n")
      cat("  ",reactant$species.name, "⇌", molesA, productA,"+", molesB, productB,"+", molesC, productC, "\n",sep=" ")
      
      cat("\nGIBBS FREE ENERGY CALCULATIONS")
      cat("\n  ΔG = Products - Reactants")
      cat("\n  ΔG = (ΔGa + ΔGb + ΔGc) - (ΔGr)\n")
      
      cat("  ΔG = (", ΔGa, "+", ΔGb, "+", ΔGc, ") - (", ΔGr, ")", sep="")
      ΔGtotal <- round(((molesA*ΔGa)+(molesB*ΔGb)+(molesC*ΔGc))-(molesR*ΔGr),3)
      cat("  ΔG =", ΔGtotal, "kJ/mol\n")
      if(ΔGtotal < 0){cat("   This process is spontaneous.\n")}
      if(ΔGtotal > 0){cat("   This process is not spontaneous.\n")}
      if(ΔGtotal == 0){cat("   This process is at equilibrium.\n")}
      return(ΔGtotal)
}
```

### The Equilibrium Constant
(1) The equilibrium constant for a reversible chemical reaction.
(2) The ratio depicting the concentrations when chemical equilibrium is reached, and expresses the extent to which equilibrium has favored in either direction.

*Note: K~sp~ describes the solubility of a substance while K~eq~ describes the state of equilibrium
```{r, message=FALSE}
equilibriumConstant <- function(solidDescription = list())
{
      ΔG <- gibbsReaction(solidDescription)
      
      logKsp <- round((-1*ΔG)/5.708,3)
      cat("\nK EQUILIBRIUM")
      cat("\n  Ksp = 10^", logKsp)
}
```

### Le Chatelier's Principle
*If a change is imposed on a system at equilibrium, the position of equilibrium will shift in a direction that tends to reduce the change*

How do we account for changes in concentration, pressure, and temperature at equilibrium?
(1) **Changes in concentration**
If a product or reactant is added to a system at equilibrium, the reaction will go inn in the direction that decreases the amount of the added constituent. If a product or reactant is removed from a system at equilibrium, the reaction will go in the direction that increases the amount of removed constituent.

(2) **Changes in pressure**
Case #1 | add or remove a gaseous product or reactant
      • If a product or reactant is added/removed to a system at equilibrium, the reaction will go inn in the direction that                     decreases/increases the amountof the added constituent.
Case #2 | add a non-reactive inert gas
      • Nothing changes, only adding more of the gas of interest would increase it's pressure system
Case #3 | change in volume
      • Decreasing the volume would reduce the number of gas molecules, increasing the volume would increase the number of gas molecules
      
(3) **Changes in temperature**
With increasing temperatures, reactions move in the direction that consumes energy (+ΔH), whereas decreasing temperatures would move in the direction that expels energy (-ΔH)

## Aqueous Complexes
*An aqueous complex is a dissolved species formed from two or more simpler species, each of which can exist in the aqueous solution* Their formation can increase the solubility of other compounds.
      
1. Considering the following reaction:  A^+^ + B^-^ ⇌ AB~(aq)~

2. We can write the equilibrium equation for this reaction in the usual way:  K~stab~ = [AB~(aq)~]/[A^+^][B^-^]

      *In this case, the equilibrium constant is called the stability constant because we are measuring the stability of         the aqueous complex (i.e., it's equilibrium with itself)*
      
3. The solubility of compounds forming an aqueous species is greater that that of the predicted solubility for the equivalent solid compound. This is because some of the ions released during the solid dissolution process are taken up by the aqueous complex:   [AB~(aq)~] = K~stab~ * [A^+^][B^-^] = K~stab~ * K~sp~
            
How does the solubility of a solid change with the formation of an aqueous complex?
```{r}
Kstab <- function(aqDescription = list())
{
      ## extract reactant G
      aqComplex <- data.frame(subcrt(aqComplex, aqComplexState, T = Temp, P = Pres))
      nameAqComplex <- paste("aqComplex$out.", aqComplex$species.name, ".G",sep ="")
      ΔGaq <- round(eval(parse(text = nameAqComplex))/1000,3)
      
      ## extract product G
      cation <- data.frame(subcrt(cation, T = Temp, P = Pres))
      nameCation <- paste("cation$out.", str_replace(cation, "[\\+\\-\\*]", "."), ".G",sep ="")
      print(nameCation[1])
      ΔGcation <- round(eval(parse(text = nameCation[1]))/1000,3)

      anion <- data.frame(subcrt(anion, T = Temp, P = Pres))
      nameAnion <- paste("anion$out.", str_replace(anion, "[\\+\\-\\*]", "."), ".G",sep ="")
      ΔGanion <- round(eval(parse(text = nameAnion[1]))/1000,3) 
      
      ## print chemical formula
      cat("REACTION DESCRIPTION")
      cat("  Reactants ⇌ Products\n")
      cat("  ",cation$species.name," + ", anion$species.name, "⇌", aqComplex$species.name, "\n",sep=" ")
      
      cat("\nGIBBS FREE ENERGY CALCULATIONS")
      cat("\n  ΔG = Products - Reactants")
      cat("\n  ΔG = (ΔGcation + ΔGanion) - (ΔGaq)\n")
      
      cat("  ΔG = (", ΔGcation, "+", ΔGanion, ") - (", ΔGaq, ")", sep="")
      ΔGtotal <- round(((molesCation*ΔGcation)+(molesAnion*ΔGanion))-(molesAq*ΔGaq),3)
      cat("  ΔG =", ΔGtotal, "kJ/mol\n")
      
      logKstab <- round((-1*ΔGtotal)/5.708,3)
      cat("\nK AQ COMPLEX STABILITY")
      cat("\n  Kstab = 10^", logKstab)
}

aqueousComplexFormation <- function(solidDescription = list(), aqDescription = list())
{
      ## equilibrium for a solid
      Keq <- equilibriumConstant(solidDescription)
      Kstab <- Kstab(aqDescription)
}
```

Describe the solid and aqueous complex of interest below.
```{r}
solidDescription = list(reactant = "gypsum",reactantState = "cr",molesR=1,
                    productA = "Ca+2",molesA=1,
                    productB = "SO4-2",molesB=1,
                    productC = "water", molesC=2,
                    Temp = 25,Pres = 1)

aqDescription = list(aqComplex = "CaSO4", aqComplexState="aq", molesAq = 1, 
                    cation = "Ca+2", molesCation = 1, 
                    anion = "SO4-2", molesAnion = 1, 
                    Temp = 25, Pres = 1)

aqueousComplexFormation(solidDescription, aqDescription)
```











