---
title: "Ch 2. Equilibrium Thermodynamics and Kinetics"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
date: '2022-06-14'
editor_options: 
  markdown: 
    wrap: 72
---
# The Laws of Thermodynamics
### First Law of Thermodynamics

*Energy cannot be created nor destroyed*

**Energy** The internal energy of a system is the sum of the kinetic and
potential energies of its constituent atoms. The *heat of formation* or
*enthalpy of formation* is the enthalpy change that occurs what a
compound is formed from its elements. *Exothermic* reactions release
heat energy and *endothermic* reactions use heat energy.

```{r, message=FALSE}
internalEnergy <- function(parameters = list(),...)
{
      # Energy change in a system by volume
      ΔE <- dq - p*dv
}
```


**Enthalpy** Enthalpy is equal to heat flow when processes occur at
constant pressure and the only work done is p-v work.

```{r, message=FALSE}
systemEnthalpy <- function(parameters = list(),...)
{
      # find ΔE
      ΔE <- internalEnergy(dq,dw,dv,p)
      
      # Enthalpy change in a system
      ΔH <- ΔE + (p*dv) + (v*dp)
}
```

### Second Law of Thermodynamics

**Entropy** *For any spontaneous process, the process always proceeds in
the direction of decreasing order.*

The release of energy is always preferable to absorption.

```{r}
systemEntropy <- function(parameters = list(),...)
{
      if(length(v)==0){
      #using ΔE
      ΔE <- internalEnergy(dq,dw,dv,p)
      ΔS <- (ΔE+(p*dv))/t
      }
      else{
      #using ΔH
      ΔH <- systemEnthalpy(dq,dw,dp,dv,p,v)
      ΔS <- (ΔH-(v*dp))/t
      }
}
```


**Gibbs Free Energy** The energy of a system approaching equilibrium is measured as *Gibbs free energy* at a constant T and P. Gibbs free energy is calculated for each species to derive the net ΔG.
```{r, message=FALSE}
GibbsFreeEnergy <- function(parameters = list(),...)
{
      ΔH <- systemEnthalpy(dq, dw, dp, dv, p, v)
      ΔS <- systemEntropy(dq,dw,dp,dv,p,v,t)
      ΔG <- ΔH - t*ΔS
}
```

Now, we can describe the system as a whole to determine whether a reaction is spontaneous.
```{r, message=FALSE}
systemConditions <- function(system = list(),...)
{
      ΔE <- internalEnergy(list(dq, dw, dv, p))
      cat("The change in energy is ", ΔE, ".", sep = "") 
      
      ΔH <- systemEnthalpy(list(dq, dw, dp, dv, p, v))
      cat("\nThe change in enthalpy is ", ΔH, ".", sep = "")
      
      ΔS <- systemEntropy(list(dq,dw,dp,dv,p,v,t))
      cat("\nThe change in entropy is ", ΔS, ".", sep = "")
      
      ΔG <- GibbsFreeEnergy(list(dq,dw,dv,dp,q,w,v,p,t))
      
      cat("\nThe change in Gibbs free energy is ", ΔG, ".", sep = "")
      if(ΔG < 0){cat("\nThus, this process is spontaneous.")}
      if(ΔG > 0){cat("\nThus, this process is not spontaneous.")}
      if(ΔG == 0){cat("\nThus, this process is at equilibrium.")}
}
```

Let's practice! Describe your system here.
```{r}
# Please replace the numbers below.
      dq <- 5           #heat change
      dw <- 3           #work change
      dp <- 4           #pressure change
      dv <- 7           #volume change
      q <- 4            #heat constant
      w <- 3            #work constant
      p <- 1            #pressure constant
      v <- 4            #volume constant
      t <- 23           #temperature constant (K)

# Execute description      
system <- list(dq, dw, dp, dv, q, w, p, v, t)
systemConditions(system)
```
