---
title: "Ch 2. Equilibrium Thermodynamics and Kinetics"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
date: '2022-06-14'
editor_options: 
  markdown: 
    wrap: 72
---
<<<<<<< HEAD

# The Laws of Thermodynamics

=======
# The Laws of Thermodynamics 
>>>>>>> f0bed09e7f58e235a69ea2ffad64f004d4437e93
### First Law of Thermodynamics

*Energy cannot be created nor destroyed*

```{r, message=FALSE}
      # define variables
      dq <- "heat change"  # added or removed
      dw <- "work change"  # done to or by the system
      dp <- "pressure change"  # via work
      dv <- "volume change"  # via work
      q <- "heat constant"
      w <- "work constant"
      p <- "pressure constant"
      v <- "volume constant"
      t <- "temperature constant (K)"
```

**Energy** The internal energy of a system is the sum of the kinetic and
potential energies of its constituent atoms. The *heat of formation* or
*enthalpy of formation* is the enthalpy change that occurs what a
compound is formed from its elements. *Exothermic* reactions release
heat energy and *endothermic* reactions use heat energy.

```{r, message=FALSE}
internalEnergy <- function(dq,dw,dv, p)
{
      # Energy change in a system
      ΔE <- dq - dw       # change in energy by work or heat
      ΔE <- dq - p*dv     # change in work by change in volume
}
```

**Enthalpy** Enthalpy is equal to heat flow when processes occur at
constant pressure and the only work done is p-v work.

```{r, message=FALSE}
systemEnthalpy <- function(dq,dw,dp,dv,p,v)
{
      # find ΔE
      ΔE <- internalEnergy(dq,dw,dv,p)
      
      # Enthalpy change in a system
      ΔH <- ΔE + (p*dv) + (v*dp)
}
```

### Second Law of Thermodynamics

**Entropy** *For any spontaneous process, the process always proceeds in
the direction of decreasing order.*

The release of energy is always preferable to absorption.

```{r}
systemEntropy <- function(dq,dw,dp,dv,p,v,t)
{
      if(length(v)==0){
      #using ΔE
      ΔE <- internalEnergy(dq,dw,dv,p)
      ΔS <- (ΔE+(p*dv))/t
      }
      else{
      #using ΔH
      ΔH <- systemEnthalpy(dq,dw,dp,dv,p,v)
      ΔS <- (ΔH-(v*dp))/t
      }
}
```

Alter input here to calculate the energy, enthalpy, and entropy of a
system.

```{r, message=FALSE}
systemConditions <- function(dq,dw,dv,dp,q,w,v,p,t)
{
      ΔE <- internalEnergy(dq, dw, dv, p)
      print(paste("The change in energy is", ΔE)) 
      
      ΔH <- systemEnthalpy(dq, dw, dp, dv, p, v)
      print(paste("The change in enthalpy is", ΔH))
      
      ΔS <- systemEntropy(dq,dw,dp,dv,p,v,t)
      print(paste("The change in entropy is", ΔS)) 
}

systemConditions(dq=5,dw=3,dv=4,dp=7,q=4,w=3,v=NULL,p=1,t=132)
```
